---
title: 'Practical Machine Learning: Course Project'
author: "YuShirota"
date: "Thursday, August 20, 2015"
output: html_document
---

# Summary
From various parameters during the exercise, the manner in which the participants did the exercise was estimated. The original dataset contains a number of NAs and apparently useless variables, so that they needed to be removed. Sixty percent of the original data were randomly extracted and served as a training set. Results of training was used to predict the manner of exercise in the varidation set, which is the other 40% of the original data, in order to estimate out-of-sample error rate.  Resulting confusion matrix showed high accuracy.

## Data loading, partition and pre-processing
The data is split into train set and validation set for cross validation.

```{r, message=FALSE}
library(caret)
pmldata <- read.csv("D:/Documents/R_Coursera/pml-training.csv",
                    header = TRUE, na.strings = "NA")
set.seed(324)
inTrain <- createDataPartition(pmldata$classe, p = 0.6, list = FALSE)
pml_train <- pmldata[inTrain, ]
pml_validation <- pmldata[-inTrain, ]
```

There are a lot of variables in the dataset.
```{r,echo=TRUE}
length(names(pml_train))
```
The dataset contains a number of useless variables. 
Here, variables where more than half are "NA" or those with near zero variance are removed. "Near zero variance" is determined using nearZeroVar function in the caret package.

```{r,echo=TRUE}
count_na <- apply(pml_train, 2, function(x) length(which(is.na(x))))
pml_train <- pml_train[, (count_na <= length(pml_train) * 0.5)]
nsv <- nearZeroVar(pml_train, saveMetrics = TRUE)
pml_train <- pml_train[, nsv$nzv == FALSE]
length(names(pml_train))
```

Now the number of variables is substantially decreased. Furthermore, several variables are just labels, i.e., subject names or recorded date/time. These are
also removed.
```{r,echo=TRUE}
pml_train <- pml_train[, 7:59]
```


## Training process
Here I choose random forest method to classify the manner in which the participants did the exercise. It is stored in the variable "classe."
```{r, message=FALSE}
modelRF <- train(classe ~., data = pml_train, method = "rf", 
                  trControl=trainControl(number=5), ntree = 100, prox = TRUE)
```

It took 689.84 seconds.

## Cross validation
Using the validation dataset, out-of-sample error rate is estimated.
```{r}
pred <- predict(modelRF, pml_validation)
confusionMatrix(pml_validation$classe, pred)$table
confusionMatrix(pml_validation$classe, pred)$overall
```
As you can see, the accuracy is high.







